---
title: "Bayesian Disease Mapping"
author: "Anthony"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Lawson Chapter 7

```{r}
  
data_jags<-list(m=46,
                Y1998=c(18,90,10,120,12,14,76,96,10,256,37,23,40,29,36,55,
                        21,63,15,19,129,47,260,60,10,184,22,45,43,44,
                        10,171,11,34,22,34,51,63,90,201,10,202,87,25,25,97),
                Exp98 = c(19.33,105.22,8.99,126.21,12.95,17.09,85.53,
                          107.178,11.0292,248.42,38.595,27.0027,
                          13.363,194.239,84.088,23.936,29.1377,
                          121.126))


```










## Lawson example Scott 
  - data obtained here https://www.paulamoraga.com/book-geospatial/sec-arealdataexamplespatial.html
  - The data contain, for each of the Scotland counties, the number of observed and expected lip cancer cases between 1975 and 1980, and a variable that indicates the proportion of the population engaged in agriculture, fishing, or forestry (AFF). The AFF variable is related to exposure to sunlight which is a risk factor for lip cancer. The data also contain a map of the Scotland counties.

```{r}
library(SpatialEpi)
data(scotland)
 class(scotland)
 names(scotland)
 
 head(scotland$data)
```
## Coordinate reference system
  - CRS is mapped to world map using geo coding.
```{r}
map <- scotland$spatial.polygon
plot(map)

codes <- rgdal::make_EPSG()
codes[which(codes$code == "27700"), ]

proj4string(map) <- "+proj=tmerc +lat_0=49 +lon_0=-2
+k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36
+units=km +no_defs"

map <- spTransform(map,
                   CRS("+proj=longlat +datum=WGS84 +no_defs"))

```

## Data preparation
  - county is the ID of each region
  - Y observed lip cancer in each county
  - E expected lip cancer in each county
  - AFF proportion engaged in agriculture, fishing and forestry
  - SIR SIR per county
  
```{r}

 d <- scotland$data[,c("county.names", "cases", "expected", "AFF")]
names(d) <- c("county", "Y", "E", "AFF")

d$SIR <- d$Y / d$E
```


